Using Fundamental Measure Theory to Treat the Correlation Function of the Inhomogeneous Hard-Sphere Fluid

### 2012-08-29 papers/contact/paper.tex
edit "S" section of theory section Tested everything on node08.

We now address the symmetrically averaged correlation function, which
is defined in Equation~\ref{eq:gS}.  This corresponds to the
correlation function averaged for spheres \emph{touching at a given
  point}.  In this case, we conceptually would like to evaluate the
pressure felt by the surface of spheres where that surface is located
at point~$\rr$.  We can approximate this value by assuming that this
pressure will be simply related to the free energy density at
point~$\rr$.  Through math similar to the previous derivations, this
leads to the expression
\begin{align}
  %% p_{HS}^S(\mathbf{r}) &= \frac{1}{n_0(\rr) 2\pi \sigma^2} \frac{\partial
  %%   a_{HS}(\rr)}{\partial \sigma} \label{eq:p_{HS}^S}\\
%  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 2\pi \sigma^2} \frac{\delta
%    A_{HS}}{\delta \sigma(\mathbf{r})} \\
  g_\sigma^S(\rr)% &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
  &= \frac{1}{n_0(\rr)^2}\frac{1}{ 2\pi \sigma^2}
  \frac{\partial \Phi(\rr)}{\partial \sigma} \label{eq:g-S}
\end{align}
where $\Phi(\rr) = \Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)$ is the
dimensionless free energy density.  This requires that we evaluate the
derivatives of the fundamental measures $n_\alpha(\rr)$ with respect
to diameter.  These derivatives, when performed exact, lead to
derivatives of the $\delta$~function, resulting in a functional gives
poor results.  We find that our $g_\sigma^S$ is considerably improved
by using integration by parts to eliminate derivatives of the
$\delta$~function by approximating the density to be smoothly varying,
which leads to derivatives such as:
\begin{align}
  \frac{\partial n_2(\rr)}{\partial \sigma} &\approx \frac{2}{\sigma}n_2(\rr)
\end{align}
We note that the density at a hard surface is \emph{not} smoothly
varying, but that the removal of these terms improves the functional,
while at the same time reducing its computational cost by avoiding the
need to calculate any additional weighted densities or convolutions.

The final formula for the symmetrically averaged correlation function
under these approximations is:
\begin{align}
  g_\sigma^S &= \text{\textcolor{red}{put formula in here, if it is
      short enough}}
\end{align}

@@

\section{Comparison with simulation}

>>>>

We now address the symmetrically averaged correlation function, which
is defined in Equation~\ref{eq:gS}.  This corresponds to the
correlation function averaged for spheres \emph{touching at a given
  point}.  In this case, we conceptually would like to evaluate the
pressure felt by the surface of spheres where that surface is located
at point~$\rr$.  We can approximate this value by assuming that this
pressure will be simply related to the free energy density at
point~$\rr$.  Through math similar to the previous derivations, this
leads to the expression
\begin{align}
  %% p_{HS}^S(\mathbf{r}) &= \frac{1}{n_0(\rr) 2\pi \sigma^2} \frac{\partial
  %%   a_{HS}(\rr)}{\partial \sigma} \label{eq:p_{HS}^S}\\
%  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 2\pi \sigma^2} \frac{\delta
%    A_{HS}}{\delta \sigma(\mathbf{r})} \\
  g_\sigma^S(\rr)% &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
  &= \frac{1}{n_0(\rr)^2}\frac{1}{ 2\pi \sigma^2}
  \frac{\partial \Phi(\rr)}{\partial \sigma} \label{eq:g-S}
\end{align}
where $\Phi(\rr) = \Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)$ is the
dimensionless free energy density.  This step is an
approximation---unlike the analogous
Equation~\ref{eq:g-A-exact}---because it assumes that we have
available a local functional $\Phi(\rr)$ whose derivative provides the
pressure needed to compute $g_\sigma(\rr)$.  Equation~\ref{eq:g-S}
requires that we evaluate the derivatives of the fundamental measures
$n_\alpha(\rr)$ with respect to diameter, which leads us to
derivatives of the $\delta$~function, which we can simplify and
approximate using an assumption of a reasonably smooth density:
\begin{align}
  \frac{\partial n_2(\rr)}{\partial \sigma}
  &= \int \delta'(\sigma - |\rr-\rr'|) n(\rr')d\rr' \\
  &= \frac2{\sigma}n_2(\rr) - \int \delta\left(\frac\sigma2 - |\rr-\rr'|\right)
  \frac{\rr-\rr'}{|\rr-\rr'|}\cdot\mathbf{\nabla}n(\rr')d\rr' \\
  &\approx \frac{2}{\sigma}n_2(\rr)
\end{align}
In the systems that we study, the density is \emph{not} reasonably
smooth, but we can state empirically making this approximation
nevertheless improves the predictions of our functional, while at the
same time reducing its computational cost by avoiding the need to
calculate any additional weighted densities or convolutions.

The final formula for our approximation to the symmetrically averaged
correlation function under these approximations is:
\begin{align}
  g_\sigma^S &= \text{\textcolor{red}{put formula in here, if it is
      short enough}}
\end{align}
We note that there is a flaw in this functional, which tracks back to
the assumption in Equation~\ref{eq:g-S} that the pressure due to
spheres touching at point~$\rr$ can be written as a derivative of
$\Phi(\rr)$.  This assumption cannot be valid in regions of space
(such as within a hard wall) where there are no contacts, but the
local energy density, as defined within FMT, is nonzero.  As we shall
see in Section~\ref{sec:comparison}, this leads to a significant and
systematic error.

@@

\section{Comparison with simulation}\label{sec:simulation}
